---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 12, message = FALSE, warning = FALSE)
library(dplyr)
library(tibble)
library(ggplot2)
library(gridExtra)

extract <- function (filename) {
  timestamp <- strsplit(filename, '.csv')[[1]]

  raw <- read.csv(paste(data.directory, filename, sep = ''))
  raw <- raw %>% mutate(date = as.Date(timestamp, '%m-%d-%Y'))
  return(raw)
}

dat <- tibble()
data.directory <- './COVID-19/csse_covid_19_data/csse_covid_19_daily_reports/'
files <- list.files(data.directory)

for (name in files) {
  if (name != 'README.md') {
    dat <- rbind(dat, extract(name))
  }
}

dat[is.na(dat)] <- 0
```

Let's look at the total number of confirmed cases of coronavirus, as far back as we have data.


```{r}
dat %>%
  group_by(date) %>%
  summarise(Confirmed = sum(Confirmed)) %>%
  ggplot() +
  aes(x = date, y = Confirmed) +
  geom_col() +
  theme_minimal() +
  labs(y = "Total Confirmed Cases", x = "Date", title = "Worldwide Total Confirmed Cases")
```

Ok that's a lot of cases, what about total number of deaths?

```{r}
dat %>%
  group_by(date) %>%
  summarise(Deaths = sum(Deaths)) %>%
  ggplot() +
  aes(x = date, y = Deaths) +
  geom_col() +
  theme_minimal() +
  labs(y = "Total Confirmed Cases", x = "Date", title = "Worldwide Total Deaths")
```

Yikes that's pretty bad. Well both of those curves begin to flatten a little near the end, does that mean that the virus is slowing down?

```{r}
newly.confirmed.cases <- dat %>%
  group_by(date) %>%
  summarise(Confirmed = sum(Confirmed)) %>%
  mutate(Confirmed = Confirmed - lag(Confirmed)) %>%
  ggplot() +
  aes(x = date, y = Confirmed) +
  geom_col() +
  theme_minimal() +
  labs(y = "Newly Confirmed Cases", x = "Date", title = "Worldwide Newly Confirmed Cases")

new.deaths <- dat %>%
  group_by(date) %>%
  summarise(Deaths = sum(Deaths)) %>%
  mutate(Deaths = Deaths - lag(Deaths)) %>%
  ggplot() +
  aes(x = date, y = Deaths) +
  geom_col() +
  theme_minimal() +
  labs(y = "Newly Confirmed Cases", x = "Date", title = "Worldwide Deaths Per Day")

grid.arrange(newly.confirmed.cases, new.deaths, ncol = 2)
```

Alright that's bad it at least it's slowing down right? Let's confirm by highlighting cases around the world relative to China.

```{r}
dat %>%
  mutate(china = ifelse(Country.Region == 'Mainland China', 'China', 'Not China')) %>%
  ggplot() +
  aes(x = date, y = Confirmed, fill = china) +
  geom_col() +
  theme_minimal() +
  labs(y = "Total Confirmed Cases", x = "Date", title = "Worldwide Total Confirmed Cases, outside of China")
```

Ok that's just barely a sliver right 😅, let's zoom in on the blue bit to ensure that we're doing all right outside China.

```{r}
dat %>%
  filter(Country.Region != 'Mainland China') %>%
  ggplot() +
  aes(x = date, y = Confirmed) +
  geom_col() +
  theme_minimal() +
  labs(y = "Total Confirmed Cases", x = "Date", title = "Worldwide Total Confirmed Cases, outside of China")
```

.... uh oh
